Automated Metadata Generation for Fish Specimen Image Collections

### 2021-03-26 paper/paper.tex
Started to edit abstract.

As a step towards improving metadata in specimen research image collections, our team has been developing methods for automatically analyzing fish images to extract a variety of important features. These fish specimens are being studied for a larger project titled Biology Guided Neural Networks (BGNN), which is developing a novel class of artificial neural networks that can exploit the machine readable and predictive knowledge about biology that is available in the form of specimen images, phylogenies and anatomy ontologies. Using a combination of machine learning and image informatics tools and techniques, we can accurately determine metadata such as fish quantity and location within images, fish orientation and other quantitative fish features, image scaling based on ruler identification and measurement, and general image quality metrics for a substantial number of the images being used in the BGNN project.

@@

%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
\begin{teaserfigure}
  \includegraphics[width=\textwidth]{fish}
  \caption{Fish prediction}
  \Description{Fish prediction}
  \label{fig:fish}
\end{teaserfigure}

>>>>

 Over the last several decades advances in computing, imaging, and cyberinfrastructure have had a major impact on scientific research and discovery. One area of considerable activity is the digitization of the biological specimens that have been acquired worldwide by museums and other research institutions; thus producing large image collections. 
Metadata is often unavailable, sparse or incorrect within these specimen image repositories, but is vital for subsequent machine learning, analysis and scientific discovery. As a step towards improving metadata in specimen research image collections, our team has been developing methods for automatically analyzing fish images to extract a variety of important features. These fish specimens are being studied for a larger project titled Biology Guided Neural Networks (BGNN), which is developing a novel class of artificial neural networks that can exploit the machine readable and predictive knowledge about biology that is available in the form of specimen images, phylogenies and anatomy ontologies. Using a combination of machine learning and image informatics tools and techniques, we can accurately determine metadata such as fish quantity and location within images, fish orientation and other quantitative fish features, image scaling based on ruler identification and measurement, and general image quality metrics for a substantial number of the images being used in the BGNN project. 
Our goal is to develop image metadata generation methods that both support the novel machine learning research underway within the BGNN project, and provide a framework for future technology developments that can be deployed by repository curators to improve and bolster the metadata they provide with their specimen images. A longer term goal is to extend the image analysis methods for computing specific quantitative features in support of specific biological investigations.

@@

%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
\begin{teaserfigure}
  \includegraphics[width=\textwidth]{fish}
  \caption{Fish prediction. Not sure such a big image should go here.  I'll think about a more succinct figure.}
  \Description{Fish prediction}
  \label{fig:fish}
\end{teaserfigure}
